Word Sense Disambiguation: Pipelining with Supervised LSA

### 2014-05-11 paper.tex
fixed merge issues

        One widely used knowledge-based algorithm is Lesk's algorithm.  Given that there are two words W1 and W2 with multiple word senses.

@@

        Another knowledge-based algorithm is semantic similarity.  In order to calculate semantic similarity, calculate path between words in Wordnet for each possible word sense of the ambiguous word with other words in the context.  To select the correct word sense, minimize the total semantic path.  The context of the word depends, and can vary from one sentence to the length of the document.  Generally, only nouns are used to provide context.

@@

	 We implemented a pipeline each with a different type of algorithm in order to do 
	 word sense disambiguation. The first step is a part-of-speech tagging step that uses
	 bigram models in order to determine what the usual part of speech is in a certain location. 
	 We use this approach for a subset of data that has high predictability.

	 Previously, we had a step involving semantic knowledge approaches, but this had low accuracy and we discarded it from the pipeline. We still detail the methods below.

 	 For the rest of the data, we applied supervised LSA, a novel implementation of the simple LSA algorithm with some ideas drawn from more sophisticated LSA approaches.

 	 We detail them in the following sections.  
 	\subsection{Part-of-Speech}
 	\subsection{Semantic Knowledge base}
        \indent We attempted a couple of methods that were dictionary-based, but they yielded accuracy rates of $ < 30\%$, so we discarded them from the pipeline.

@@

	We also tried a slightly different approach to capture more of the information: For each word, and for each sense of the word, we created a vector by applying cosine similarity to each eigentopic column vector for our training vector (created from paragraph). We averaged over all the training data for each word sense. Then for testing, we created this vector from the test paragraph and compared via cosine similarity again to determine which word sense was most likely. 
	This approach had issues and we got very bad results. 
	It might be better in the future to try a modified approach to this (i.e. do something other than average the vectors).

	\section{Results}
	From Senseval-2:

@@

	Supervised LSA: 
	a) With only the Brown Corpus categories as column vectors ($15$ column vectors -> reduced to $9$) in the term-document matrix: ~ $42\%$
	b) With the Brown Corpus categories + the word-sense categories (in total, 861 column vectors -> reduced to $140$ column vectors after SVD): ~ $38\%$
	-- highly dependent on the training data -- different results for different words, some words had lower percentages, some had higher percentages (as low as $7\%$, as high as $61\%$)
	-- easier to do ambiguous words with fewer senses

>>>>

        One widely used knowledge-based algorithm is Lesk's algorithm.  The algorithm is as follows:

        Given that there are two words W1 and W2 with multiple word senses:

@@

        Another knowledge-based algorithm is semantic similarity.  In order to calculate semantic similarity, calculate the path between words in Wordnet for each possible word sense of the ambiguous word with other words in the context.  To select the correct word sense, minimize the total semantic path.  The context of the word can vary from one sentence to the length of the document.  Generally, only nouns are used to provide context.

@@

	 Our approach was essentially a pipeline of different sorts of models. 
	 The first stage consisted of part-of-speech tagging and bigram models, which
	 was used for the subset of data that had high predictability. 
	 Originally, we implemented a step that included knowledge-base approaches like
	 semantic similarity, but later discarded it from the pipeline because its accuracy
	 was not very good.

	 For the rest of the data, we implemented a Supervised LSA approach, which was to
	 an extent a novel approach.

	 We detail the methods in the following sections.  
 	\subsection{Part-of-Speech}
 	\subsection{Semantic Knowledge base}
        \indent We attempted a couple of methods that were dictionary-based, but they yielded accuracy rates of $ < 30\%$, so we discarded them from the pipeline.

@@

	We also tried a slightly different approach to capture more of the information: For each word, and for each sense of the word, we created a vector by applying cosine similarity to each eigentopic column vector for our training vector (created from paragraph). We averaged over all the training data for each word sense. Then for testing, we created this vector from the test paragraph and compared via cosine similarity again to determine which word sense was most likely. 
	This approach had issues and we got very bad results. 
	It might be better in the future to try a modified approach to this (i.e. do something other than average the vectors).
	\section{Results}
	From Senseval-2:

@@

	\subsection{Supervised LSA}

	a) With only the Brown Corpus categories as column vectors ($15$ column vectors $\rightarrow$ reduced to $9$) in the term-document matrix: $ ~ 42\%$
	b) With the Brown Corpus categories + the word-sense categories (in total, 861 column vectors $\rightarrow$ reduced to $140$ column vectors after SVD): $ ~ 38\%$
	-- highly dependent on the training data -- different results for different words, some words had lower percentages, some had higher percentages (as low as $7\%$, as high as $61\%$)
	-- easier to do ambiguous words with fewer senses
